### 2018-05-23 Datelife_paper1.Rmd
something

In the past decades, the possibility to obtain DNA sequences in large quantities from a wide variety of organisms became a reality, which, coupled to methodological development in phylogenetic and dating inference, allowed the application of molecular dating methods on a very large amount and diversity of organisms, greatly increasing the quantity of<!--organisms with--> data on taxon ages across the tree of life.
To date, there is a large amount of both fossil and molecular-based data on taxon ages and phylogenetic relationships in public repositories such as Dryad, TreeBASE and Open Tree of Life (OToL).
OToL alone holds more than 200 chronograms. <!--cache updated jan 15 2018
devtools::load_all(pkg = "~/Desktop/datelife")
length(opentree_chronograms$trees)
names(opentree_chronograms)
-->
<!--How many trees? How many dated trees?-->
Methods to include living and fossil lineages are in continued development and increased usage by the community<!--total evidence and revbayes (fossilized birth-death)-->, which coupled to better sharing data practices, are greatly contributing to the accumulation in number and type of available data on taxon ages.

@@

Total ssp number estimated to be between 3 and 100 million species, as mentiones in [@Mora2011]:
ttol_spp <- 97085
(ttol_spp * 100) / 3e+6  # 3.236167
(ttol_spp * 100) / 100e+6  # 0.097085
-->
but is currently one of the most important sources on divergence times, since it includes all main lineages of life (plants, animals, fungi, bacteria and archaea) and is currenty the largest chronogram in existence. As a science communication tool it is very powerful, since it has a friendly graphical interface, with informative while colorful outputs, that allows the general public to satisfy curiosity regarding their favorite organism or group of them. However, the primary data source and the synthetic chornogram are not publicly available, and large-scale, automated, data-mining is not allowed, minimizing its potential usefulness for the scientific community.
<!--
While its node-time search is very powerful, using available data to infer divergence time from lineages that have never been sequenced before, its timetree option is less powerful, their main flaw being that it does not provide source data.
However, their primary database is not fully open, giving information on a per node basis. So, it is not possible to easily access information from original studies for a group of taxa (One would need to perform a pair of taxa search for all pairs from the group of interest).
Divergence time estimates for all taxa in the TToL are precomputed and stored in the TimeTree database
All searches allow species names and higher group names. Common names are allowed in all searches but a list of taxa.
one-taxon search was nor working for dog, though.
1) ttol one lineage: it gives the evolutionary timeline of a lineage, showing the age of main divergence events occuring along the lineage evolutionary history, i.e., when it shared a comon ancestor with all main lineages from which it diverged a long time ago.
It shows results only graphically on the website.
2) ttol parwise search: allows common names, species names and higher group names
    how it handles names missing from tree: it looks in the NCBI taxonomy for the closest relative of the species or higher taxa. Then uses the closest relative to get the lineage pair age from ttol.
3) ttol group of taxa
4) ttol list of taxa: Cannot get it to work; does not work with cat, dog, turtle search, felidae, hominidae, phrynops search, some_ants species name search. What kind of files does it accepts: txt, rtf, doc, docx?
-->

>>>>

In the past decades, the possibility to obtain DNA sequences in large quantities from a wide variety of organisms became a reality, which, coupled to methodological development in phylogenetic and dating inference, allowed the application of molecular dating methods on a very large amount and diversity of organisms, greatly increasing the quantity of<!--organisms with--> data on taxon ages across the tree of life.
<!-- figure showing growth on studies with age data from references table from ttol-->
To date, there is a large amount of both fossil and molecular-based data on taxon ages and phylogenetic relationships in public repositories such as Dryad, TreeBASE and Open Tree of Life (OToL).
OToL alone holds more than 200 chronograms. <!--cache updated jan 15 2018
devtools::load_all(pkg = "~/Desktop/datelife")
length(opentree_chronograms$trees)
names(opentree_chronograms)
-->
<!--How many trees? How many dated trees?-->
Methods to include living and fossil lineages are in continued development and increased usage by the community<!--total evidence and revbayes (fossilized birth-death)-->, which coupled to better sharing data practices, are greatly contributing to the accumulation in number and type of available data on taxon ages.

@@

Total ssp number estimated to be between 3 and 100 million species, as mentiones in [@Mora2011]:
ttol_spp <- 97085
(ttol_spp * 100) / 3e+6  # 3.236167
(ttol_spp * 100) / 100e+6  # 0.097085
-->
but is currently one of the most important sources on divergence times, since it includes all main lineages of life (plants, animals, fungi, bacteria and archaea) and is currenty the largest chronogram in existence. As a science communication tool it is very powerful, since it has a friendly graphical interface, with informative while colorful outputs, that allows the general public to satisfy curiosity regarding their favorite organism or group of them. However, the primary source of data and the synthetic chronogram are not publicly available, and large-scale, automated, data-mining is not allowed, minimizing its potential usefulness for the scientific community.
<!--
While its node-time search is very powerful, using available data to infer divergence time from lineages that have never been sequenced before, its timetree option is less powerful, their main flaw being that it does not provide source data.
However, their primary database is not fully open, giving information on a per node basis. So, it is not possible to easily access information from original studies for a group of taxa (One would need to perform a pair of taxa search for all pairs from the group of interest).
Divergence time estimates for all taxa in the TToL are precomputed and stored in the TimeTree database
All searches allow species names and higher group names. Common names are allowed in all searches but a list of taxa.
one-taxon search was nor working for dog, though.
1) ttol one lineage: it gives the evolutionary timeline of a lineage, showing the age of main divergence events occuring along the lineage evolutionary history, i.e., when it shared a comon ancestor with all main lineages from which it diverged a long time ago.
It shows results only graphically on the website.
2) ttol parwise search: allows common names, species names and higher group names
    how it handles names missing from tree: it looks in the NCBI taxonomy for the closest relative of the species or higher taxa. Then uses the closest relative to get the lineage pair age from ttol.
3) ttol group of taxa
4) ttol list of taxa: Cannot get it to work; does not work with cat, dog, turtle search, felidae, hominidae, phrynops search, some_ants species name search. What kind of files does it accepts: txt, rtf, doc, docx?
-->
